---
title: "ColBioProject_ZIMMERMAN_FRENZE_STONEFIELD_LANG_URBANOWICZ_MORALES"
format: html
editor: visual
---

## Load Packages

```{r}
# Everyone as needed:
library(tidyverse)
```

## Load Data

```{r}
# Ian
# Pull data from https://github.com/Stoney044/Ebird_IPEDS

EB_url <- "https://raw.githubusercontent.com/Stoney044/Ebird_IPEDS/main/EBbird.csv"
IP_url <- "https://raw.githubusercontent.com/Stoney044/Ebird_IPEDS/main/IPEDS.csv"

# Load the data
Ebird <- read_csv(EB_url)
IPEDS <- read_csv(IP_url)

# Clean Metadata
# Ian
metadata <- metadata |>
  select(varTitle, varName) 

IPEDS_columns <- IPEDS |> 
  colnames()

# Filter the metadata only to include the columns in the IPEDS data
metadata <- metadata |>
  filter(varName %in% IPEDS_columns)

# export metadata as csv file 
# write_csv(metadata, "Data/Clean_IPEDS_metadata.csv")
```

## Bird Species Richness per Campus (2)

### Richness Per Sampling Event

```{r}
# Ingrid, Lee, Elizabeth, Jake, Ian

species_richness <- Ebird |> 
  group_by(UNITID, `SAMPLING EVENT IDENTIFIER`) |> 
  mutate(species_count = n_distinct(`SCIENTIFIC NAME`))

```

### Average Sampling Event Richness per UNITID

```{r}
# Ingrid, Lee, Elizabeth, Jake, Ian

mean_species_richness <- species_richness |> 
  group_by(UNITID) |> 
  summarise(mean_species_richness = mean(species_count, na.rm = TRUE))

```

## Join Species Richness to ipeds data (3)

```{r}
# Elizabeth

merged_data <- IPEDS |> 
  left_join(mean_species_richness)

```

## Explore bird species richness and impervious surface (4)

### Hypothesis (4a)

There will be less richness with more impervious surface because these surfaces cannot provide the resources birds need to survive.

### Relationship Across Colleges (4b)

```{r}
# Lee




merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = mean_impervious, y = mean_species_richness)) +
  geom_smooth(se = FALSE) + labs(x = "Mean Impervious Surface", y = "Mean Species Richness")
  
merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = log(mean_impervious), y = mean_species_richness)) +
  geom_smooth(se = FALSE) + labs(x = "Log(Mean Impervious Surface)", y = "Mean Species Richness")  

merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = mean_impervious, y = log(mean_species_richness))) +
  geom_smooth(se = FALSE) + labs(x = "Mean Impervious Surface", y = "Log(Mean Species Richness)")

merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = log(mean_impervious), y = log(mean_species_richness))) +
  geom_smooth(se = FALSE) + labs(x = "Log(Mean Impervious Surface", y = "Log(Mean Species Richness)")

```

### Mechanism (4c)

Mostly stable then dips, implies that there is a threshold of impervious surface that is okay before it causes problems.

### By Taxonomic Order (4d)


```{r}
# Lee

Ebird |> 
  group_by(UNITID, ORDER) |> 
  mutate(species_count = n_distinct(`SCIENTIFIC NAME`)) |> 
  group_by(UNITID, ORDER) |> 
  summarise(mean_species_richness = mean(species_count, na.rm = TRUE)) |> 
  right_join(IPEDS) |> 
  ggplot(aes(x = mean_impervious, y = mean_species_richness)) +
  geom_smooth(se = FALSE) +
  facet_wrap(~ORDER, scales = "free_y")

```

## Team Member Individual Explorations

### Zimmerman

### Frenze
```{r}
merged_data_NF<-merged_data|>select(
  INSTNM, INSTSIZE, ENRTOT, mean_impervious, mean_species_richness)|>
  filter(!is.na(mean_species_richness))|>
  filter(!is.na(ENRTOT))|>
  mutate(INSTSIZE = factor(INSTSIZE))

Enrollment_range<-c("Under 1,000", "1,000-5,000", "5,000-10,000", "10,000-20,000", "20,000+")

plot_1NF<-ggplot(data = merged_data_NF) +
  geom_boxplot(aes(x = INSTSIZE, y = mean_species_richness, color = INSTSIZE)) +
  labs(x = "Institutation Size", y = "Mean Species Richness", color = "Number of Students Enrolled") + 
  theme(legend.position = "bottom") +
  scale_color_discrete(labels = Enrollment_range)  

species_richness_per_enrollment<-merged_data_NF|>mutate(
  Species_richness_per_student = mean_species_richness/ENRTOT,
  .keep = "all"
)

plot_2NF<-ggplot(data = species_richness_per_enrollment) +
  geom_boxplot(aes(x = INSTSIZE, y = Species_richness_per_student, color = INSTSIZE)) +
  labs(x = "Institutation Size", y = "Mean Species Richness per Student", color = "Number of Students Enrolled") + 
  theme(legend.position = "none") +
  scale_color_discrete(labels = Enrollment_range) 
print(plot_1NF + plot_2NF)
```

There is no apparent correlation between the average impervious surface or institution size. This data set suggest there is little correlation between institution size and species richness. \# However it should be noted that seeing that even the smaller institution sizes have equitable species richness the larger ones which would logically require a larger campuses on average. This could indicate that smaller institutions may have higher species richness per square mile and with additional data pertaining to the each individual campuses area further analysis can be done.
### Stonefield

### Lang

### Urbanowicz

### Morales
