---
title: "ColBioProject_ZIMMERMAN_FRENZE_STONEFIELD_LANG_URBANOWICZ_MORALES"
format: html
editor: visual
---

## Load Packages

```{r}
# Everyone as needed:
library(tidyverse)
```

## Load Data

```{r}
# Ian
# Pull data from https://github.com/Stoney044/Ebird_IPEDS

EB_url <- "https://raw.githubusercontent.com/Stoney044/Ebird_IPEDS/main/EBbird.csv"
IP_url <- "https://raw.githubusercontent.com/Stoney044/Ebird_IPEDS/main/IPEDS.csv"

# Load the data
Ebird <- read_csv(EB_url)
IPEDS <- read_csv(IP_url)
```


## Bird Species Richness per Campus (2)
species_richness <- Ebird %>%
+     group_by(UNITID, `SAMPLING EVENT IDENTIFIER`) %>%
+     summarise(species_count = n_distinct(`SCIENTIFIC NAME`))
mean_species_richness <- species_richness %>%
+     group_by(UNITID) %>%
+     summarise(mean_species_richness = mean(species_count, na.rm = TRUE))

## Join Species Richness to ipeds data (3)
merged_data <- IPEDS %>%
+     left_join(mean_species_richness, by = "UNITID") 

## Explore bird species richness and impervious surface (4)

### Hypothesis (4a)

### Relationship Across Colleges (4b)

### Mechanism (4c)

### By Taxonomic Order (4d)

## Team Member Individual Explorations

### Zimmerman

### Frenze

### Stonefield

### Lang

### Urbanowicz

### Morales
