---
title: "ColBioProject_ZIMMERMAN_FRENZE_STONEFIELD_LANG_URBANOWICZ_MORALES"
format: html
editor: visual
---

## Load Packages

```{r}
# Everyone as needed:
library(tidyverse)
```

## Load Data

```{r}
# Ian
# Pull data from https://github.com/Stoney044/Ebird_IPEDS

EB_url <- "https://raw.githubusercontent.com/Stoney044/Ebird_IPEDS/main/EBbird.csv"
IP_url <- "https://raw.githubusercontent.com/Stoney044/Ebird_IPEDS/main/IPEDS.csv"

# Load the data
Ebird <- read_csv(EB_url)
IPEDS <- read_csv(IP_url)

# Clean Metadata
# Ian
metadata <- metadata |>
  select(varTitle, varName) 

IPEDS_columns <- IPEDS |> 
  colnames()

# Filter the metadata only to include the columns in the IPEDS data
metadata <- metadata |>
  filter(varName %in% IPEDS_columns)

# export metadata as csv file 
# write_csv(metadata, "Data/Clean_IPEDS_metadata.csv")
```

## Bird Species Richness per Campus (2)

### Richness Per Sampling Event

```{r}
# Ingrid, Lee, Elizabeth, Jake, Ian

species_richness <- Ebird |> 
  group_by(UNITID, `SAMPLING EVENT IDENTIFIER`) |> 
  mutate(species_count = n_distinct(`SCIENTIFIC NAME`))

```

### Average Sampling Event Richness per UNITID

```{r}
# Ingrid, Lee, Elizabeth, Jake, Ian

mean_species_richness <- species_richness |> 
  group_by(UNITID) |> 
  summarise(mean_species_richness = mean(species_count, na.rm = TRUE))

```

## Join Species Richness to ipeds data (3)

```{r}
# Elizabeth

merged_data <- IPEDS |> 
  left_join(mean_species_richness)

```

## Explore bird species richness and impervious surface (4)

### Hypothesis (4a)

There will be less richness with more impervious surface because these surfaces cannot provide the resources birds need to survive.

### Relationship Across Colleges (4b)

```{r}
# Lee

merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = mean_impervious, y = mean_species_richness)) +
  geom_smooth(se = FALSE)
  
merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = log(mean_impervious), y = mean_species_richness)) +
  geom_smooth(se = FALSE)  

merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = mean_impervious, y = log(mean_species_richness))) +
  geom_smooth(se = FALSE)

merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = log(mean_impervious), y = log(mean_species_richness))) +
  geom_smooth(se = FALSE)

```

### Mechanism (4c)

Mostly stable then dips, implies that there is a threshold of impervious surface that is okay before it causes problems.

### By Taxonomic Order (4d)


```{r}
# Lee

Ebird |> 
  group_by(UNITID, ORDER) |> 
  mutate(species_count = n_distinct(`SCIENTIFIC NAME`)) |> 
  group_by(UNITID, ORDER) |> 
  summarise(mean_species_richness = mean(species_count, na.rm = TRUE)) |> 
  right_join(IPEDS) |> 
  ggplot(aes(x = mean_impervious, y = mean_species_richness)) +
  geom_smooth(se = FALSE) +
  facet_wrap(~ORDER, scales = "free_y")

```

## Team Member Individual Explorations

### Zimmerman

### Frenze

### Stonefield

land_grant_SR <- merged_data |> 
  select(UNITID, INSTNM, mean_species_richness, LANDGRNT)|>
  filter(is.na(mean_species_richness) == FALSE) |>
  mutate(LANDGRNT = ifelse(LANDGRNT == 1, "Land Grant", "Not Land Grant"))

land_grant_SR |>
  ggplot(aes(x = LANDGRNT, y = mean_species_richness)) +
  geom_boxplot(aes(fill = LANDGRNT)) +
  labs(title = "Mean Species Richness of Land Grant Universities",
       x = "Land Grant Status",
       y = "Mean Species Richness") +
  theme_minimal() +
  scale_fill_manual(values = c("Land Grant" = "green", "Not Land Grant" = "skyblue2"))

### Lang

### Urbanowicz

### Morales
