---
title: "ColBioProject_ZIMMERMAN_FRENZE_STONEFIELD_LANG_URBANOWICZ_MORALES"
format: html
editor: visual
---

## Load Packages

```{r}
# Everyone as needed:
library(tidyverse)
```

## Load Data

```{r}
# Ian
# Pull data from https://github.com/Stoney044/Ebird_IPEDS

EB_url <- "https://raw.githubusercontent.com/Stoney044/Ebird_IPEDS/main/EBbird.csv"
IP_url <- "https://raw.githubusercontent.com/Stoney044/Ebird_IPEDS/main/IPEDS.csv"

# Load the data
Ebird <- read_csv(EB_url)
IPEDS <- read_csv(IP_url)

# Clean Metadata
# Ian
metadata <- metadata |>
  select(varTitle, varName) 

IPEDS_columns <- IPEDS |> 
  colnames()

# Filter the metadata only to include the columns in the IPEDS data
metadata <- metadata |>
  filter(varName %in% IPEDS_columns)

# export metadata as csv file 
# write_csv(metadata, "Data/Clean_IPEDS_metadata.csv")
```

## Bird Species Richness per Campus (2)

### Richness Per Sampling Event

```{r}
# Ingrid, Lee, Elizabeth, Jake, Ian

species_richness <- Ebird |> 
  group_by(UNITID, `SAMPLING EVENT IDENTIFIER`) |> 
  mutate(species_count = n_distinct(`SCIENTIFIC NAME`))

```

### Average Sampling Event Richness per UNITID

```{r}
# Ingrid, Lee, Elizabeth, Jake, Ian

mean_species_richness <- species_richness |> 
  group_by(UNITID) |> 
  summarise(mean_species_richness = mean(species_count, na.rm = TRUE))

```

## Join Species Richness to ipeds data (3)

```{r}
# Elizabeth

merged_data <- IPEDS |> 
  left_join(mean_species_richness)

```

## Explore bird species richness and impervious surface (4)

### Hypothesis (4a)

There will be less richness with more impervious surface because these surfaces cannot provide the resources birds need to survive.

### Relationship Across Colleges (4b)

```{r}
# Lee

merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = mean_impervious, y = mean_species_richness)) +
  geom_smooth(se = FALSE)
  
merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = log(mean_impervious), y = mean_species_richness)) +
  geom_smooth(se = FALSE)  

merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = mean_impervious, y = log(mean_species_richness))) +
  geom_smooth(se = FALSE)

merged_data |>
  group_by(UNITID) |> 
  ggplot(aes(x = log(mean_impervious), y = log(mean_species_richness))) +
  geom_smooth(se = FALSE)

```

### Mechanism (4c)

Mostly stable then dips, implies that there is a threshold of impervious surface that is okay before it causes problems.

### By Taxonomic Order (4d)


```{r}
# Lee

Ebird |> 
  group_by(UNITID, ORDER) |> 
  mutate(species_count = n_distinct(`SCIENTIFIC NAME`)) |> 
  group_by(UNITID, ORDER) |> 
  summarise(mean_species_richness = mean(species_count, na.rm = TRUE)) |> 
  right_join(IPEDS) |> 
  ggplot(aes(x = mean_impervious, y = mean_species_richness)) +
  geom_smooth(se = FALSE) +
  facet_wrap(~ORDER, scales = "free_y")

```

## Team Member Individual Explorations

### Zimmerman

### Frenze

### Stonefield

### Lang

### Urbanowicz
#Variable: LOCALE
> filtered_data <- merged_data |>
+     drop_na(LOCALE, mean_species_richness)
> ggplot(filtered_data, aes(x = LOCALE, y = mean_species_richness)) +
+     geom_jitter(width = 0.3, height = 0, alpha = 0.6) +  
+     labs(x = "Locale", y = "Bird Species Richness", title = "Bird Species Richness by Urbanization") +
+     theme_minimal()

#What is the hypothesis that you are testing? What is the reason that it makes sense to look at the hypothesis?
As the degree of urbanization increases, bird species richness decreases resulting in a negative relationship. It is important to look at the effects of urbanization in terms of bird species richness because of its drastic habitat alteration, negative impact on bird species and the resulting ecological responses, and research implications. An increase in LOCALE may result in situations such as deforestation and fragmentation of local habitats which leads to a loss of suitable habitats for different bird species. Urbanization may also reduce different resources for birds such as nesting sites and food; however, even if certain birds may be able to adapt to human development, it may lead to competition with invasive species lessening the likelihood of a diverse bird population. Studying different levels of urbanization may also give insight into human environmental changes and its effects on wildlife which could be helpful in terms of conservation efforts. Finally, it could contribute to different research which could help in creating policies for conservation and a more sustainable way of building urban environments. 
#Does accounting for impervious surface or taxonomic order change your results
Accounting for impervious surfaces may help control for the effects of habitat alteration and urban development on bird species richness and has the ability to provide a better understanding of the impact of urbanization on bird diversity. In regards to taxonomic order, it can reveal how different bird groups respond to urbanization, specifically how it can help identify specific groups that are more sensitive or tolerant to urbanization pressures.


#In regards to impervious surfaces:
ggplot(filtered_data, aes(x = LOCALE, y = mean_species_richness, color = mean_impervious)) +
+     geom_jitter(width = 0.3, height = 0, alpha = 0.6) +  
+     labs(x = "Locale", y = "Bird Species Richness", title = "Bird Species Richness by Urbanization") +
+     theme_minimal()

#In regards to taxonomic order:
> ordermerged_data <- merge(filtered_data, Ebird, by = "UNITID", all.x = TRUE)
> ggplot(ordermerged_data, aes(x = LOCALE, y = mean_species_richness)) +
+     geom_jitter(width = 0.3, height = 0, alpha = 0.6) +  
+     labs(x = "Locale", y = "Bird Species Richness", title = "Bird Species Richness by Urbanization") +
+     theme_minimal() +
+     facet_wrap(~ ORDER, scales = "free_y", ncol = 2)

### Morales
